---
title: "Performing ANOVA"
format: 
  html:
    toc: true
    html-math-method: mathjax
vignette: >
  %\VignetteIndexEntry{Performing ANOVA}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(flipr)
library(ggplot2)
load("../R/sysdata.rda")

pvalue1 <- df_1anova$pvalue
pvalue1_est <- df_1anova$pvalue_est
point_estimate1 <- df_1anova$point_estimate
grid1 <- df_1anova$grid

pvalue2 <- df_2anova$pvalue
pvalue2_est <- df_2anova$pvalue_est
point_estimate2 <- df_2anova$point_estimate
grid2 <- df_2anova$grid
```

The package allows user to perform (M)-ANOVA. Two examples are shown in this
article for more clarity. User can pass data in the plausibility function as two
possible formats:

- An object with all samples and a factor containing group memberships.
- Each sample representing a group.

Both cases will appear in this article. Since we use Fisher tests, it is 
important to change the `alternative` parameter of the plausibility function to
`right_tail`.

## Plausibility function for 1-ANOVA ðŸ“

For this example, data represent chicken weights by feed type. There are three
populations that have been fed different types of food. The null hypothesis is 

$$
H_0: \mu_1 = \mu_2 = \mu_3
$$

with $\mu_i$ representing the sample mean in the group $i$.

```{r, eval=FALSE}
data <- chickwts$weight[1:36] # 1 variable
memberships <- as.factor(chickwts$feed[1:36]) # 1 factor with 3 populations

# takes all samples except the first one
null_spec <- function(l, parameters) {
  purrr::map2(l, parameters, \(l, param) l - param)
}
stat_functions <- list(stat_anova_f)

# all means except first one
stat_assignments <- list(mu2 = 1, mu3 = 1)

pf <- PlausibilityFunction$new(
  null_spec = null_spec,
  stat_functions = stat_functions,
  stat_assignments = stat_assignments,
  data, memberships
)
pf$set_alternative("right_tail")

pvalue1 <- pf$get_value(rep(0, 2))
```

```{r}
pvalue1
```

```{r, eval=FALSE}
samples <- purrr::map(unique(as.numeric(memberships)), \(.class) {
  data[which(as.numeric(memberships) == .class)]
})
means <- c(mean(samples[[2]]) - mean(samples[[1]]),
           mean(samples[[3]]) - mean(samples[[1]]))
pvalue1_est <- pf$get_value(means)
```

```{r}
pvalue1_est
```

```{r, eval=FALSE}
pf$set_point_estimate(means)
point_estimate1 <- pf$point_estimate
```

```{r}
point_estimate1
```

```{r, eval=FALSE}
pf$set_parameter_bounds(
  point_estimate = pf$point_estimate,
  conf_level = pf$max_conf_level
)

pf$set_grid(
  parameters = pf$parameters,
  npoints = 10
)
pf$evaluate_grid(grid = pf$grid)
grid1 <- pf$grid
```

```{r}
grid1 %>% 
  ggplot(aes(mu2, mu3, z = pvalue)) + 
  geom_contour_filled(binwidth = 0.05) + 
  labs(
    title = "Contour plot of the p-value surface", 
    subtitle = "Using Fisher's non-parametric combination", 
    x = expression(paste(mu, "_2")), 
    y = expression(paste(mu, "_3")), 
    fill = "p-value"
  ) + 
  theme_minimal()
```


## Plausibility function for 2-ANOVA ðŸŒ¸

In this example, data used represent flower caracteristics by their specie. The 
variables are sepal length and sepal width that are compared bewteen three 
different flower species. The null hypothesis is 

$$
H_0: \begin{pmatrix}
\mu_{11} \\
\mu_{12}
\end{pmatrix}
= 
\begin{pmatrix}
\mu_{21} \\
\mu_{22}
\end{pmatrix}
= 
\begin{pmatrix}
\mu_{31} \\
\mu_{32}
\end{pmatrix}
$$

with $\mu_{ij}$ representing the sample mean in the group $i$ for the variable 
$j$.

```{r, eval=FALSE}
iris_setosa <- iris |> filter(Species == "setosa")
iris_setosa <- iris_setosa[1:10,1:2]
iris_versicolor <- iris |> filter(Species == "versicolor")
iris_versicolor <- iris_versicolor[1:10,1:2]
iris_virginica <- iris |> filter(Species == "virginica")
iris_virginica <- iris_virginica[1:10,1:2]

# takes all samples except the first one
null_spec <- function(l, parameters) {
  parameters <- split(parameters, ceiling(seq_along(parameters) / (length(parameters)/length(l))))
  purrr::map2(l, parameters, \(group, params) {
    purrr::map2(group, params, \(var, param) {
      var - param
    }) |> as.data.frame()
  })
}
stat_functions <- list(stat_anova_f)

# all means except ones from first sample
stat_assignments <- list(
  mu2_1 = 1, mu2_2 = 1,
  mu3_1 = 1, mu3_2 = 1)

pf <- PlausibilityFunction$new(
  null_spec = null_spec,
  stat_functions = stat_functions,
  stat_assignments = stat_assignments,
  iris_setosa, iris_versicolor, iris_virginica
)
pf$set_alternative("right_tail")

pvalue2 <- pf$get_value(rep(0, 4))
```

```{r}
pvalue2
```

```{r, eval=FALSE}
d1 <- c(
  mean(iris_versicolor$Sepal.Length) - mean(iris_setosa$Sepal.Length),
  mean(iris_versicolor$Sepal.Width)  - mean(iris_setosa$Sepal.Width)
)
d2 <- c(
  mean(iris_virginica$Sepal.Length) - mean(iris_setosa$Sepal.Length),
  mean(iris_virginica$Sepal.Width)  - mean(iris_setosa$Sepal.Width)
)
pvalue1_est <- pf$get_value(c(d1, d2))
```

```{r}
pvalue1_est
```

```{r, eval=FALSE}
pf$set_point_estimate(c(d1, d2))
point_estimate2 <- pf$point_estimate
```

```{r}
point_estimate2
```

```{r, eval=FALSE}
pf$set_parameter_bounds(
  point_estimate = pf$point_estimate,
  conf_level = pf$max_conf_level
)

pf$set_grid(
  parameters = pf$parameters,
  npoints = 10
)
pf$evaluate_grid(grid = pf$grid)

grid2 <- pf$grid
```

```{r}
grid2 |> dplyr::filter(pvalue > 0.99)
```


## Conclusion

For these two examples, means are different in each population. Thus, when we 
test the equality of means with `get_value(0)`, we get a pvalue close to 0, 
allowing us to reject the hypothesis that the means are equal within the 
different populations. On the other hand, if we get the pvalue for the test on 
the estimator values, it is equal to 1, the largest possible value, which is 
coherent.




