---
title: "Performing ANOVA"
format: 
  html:
    toc: true
    html-math-method: mathjax
vignette: >
  %\VignetteIndexEntry{Performing ANOVA}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(flipr)
library(ggplot2)
load("../R/sysdata.rda")

pvalue1 <- df_1anova$pvalue
point_estimate1 <- df_1anova$point_estimate
grid1 <- df_1anova$grid

pvalue2 <- df_2anova$pvalue
point_estimate2 <- df_2anova$point_estimate
grid2 <- df_2anova$grid
```

The package allows user to perform (M)-ANOVA. Two examples are shown in this
article for more clarity. User can pass data in the plausibility function as two
possible format:

- An object with all samples and a factor containing group memberships.
- Each sample representing a group.

Both cases will appear in this article.

## Plausibility function for 1-ANOVA

For this example, data represent chicken weights by feed type. The null 
hypothesis is 
$$
H_0: \mu_1 = \mu_2 = \mu_3
$$
with $\mu_i$ representing the sample mean in the group $i$.

```{r, eval=FALSE}
data <- chickwts$weight[1:36] # 1 variable
memberships <- as.factor(chickwts$feed[1:36]) # 1 factor with 3 populations

# takes all samples except the first one
null_spec <- function(l, parameters) {
  purrr::map2(l, parameters, \(l, param) l - param)
}
stat_functions <- list(stat_anova_f)

# all means except first one
stat_assignments <- list(mu2 = 1, mu3 = 1)

pf <- PlausibilityFunction$new(
  null_spec = null_spec,
  stat_functions = stat_functions,
  stat_assignments = stat_assignments,
  data, memberships
)

pvalue1 <- pf$get_value(rep(0, 2))
```

```{r}
pvalue1
```

```{r, eval=FALSE}
samples <- purrr::map(unique(as.numeric(memberships)), \(.class) {
  data[which(as.numeric(memberships) == .class)]
})
means <- c(mean(samples[[2]]) - mean(samples[[1]]),
           mean(samples[[3]]) - mean(samples[[1]]))
pf$set_point_estimate(means)
point_estimate1 <- pf$point_estimate
```

```{r}
point_estimate1
```

```{r, eval=FALSE}
pf$set_parameter_bounds(
  point_estimate = pf$point_estimate,
  conf_level = pf$max_conf_level
)

pf$set_grid(
  parameters = pf$parameters,
  npoints = 10
)
pf$evaluate_grid(grid = pf$grid)
grid1 <- pf$grid
```

```{r}
grid1 %>% 
  ggplot(aes(mu2, mu3, z = pvalue)) + 
  geom_contour_filled(binwidth = 0.05) + 
  labs(
    title = "Contour plot of the p-value surface", 
    subtitle = "Using Fisher's non-parametric combination", 
    x = expression(paste(mu, "_2")), 
    y = expression(paste(mu, "_3")), 
    fill = "p-value"
  ) + 
  theme_minimal()
```


## Plausibility function for 2-ANOVA

In this example, data used represent flower caracteristics by their specie.
The null hypothesis is 
$$
H_0: \begin{pmatrix}
\mu_{11} \\
\mu_{12}
\end{pmatrix}
= 
\begin{pmatrix}
\mu_{21} \\
\mu_{22}
\end{pmatrix}
= 
\begin{pmatrix}
\mu_{31} \\
\mu_{32}
\end{pmatrix}
$$
with $\mu_{ij}$ representing the sample mean in the group $i$ for the variable 
$j$.

```{r, eval=FALSE}
n <- 10
x1 <- rnorm(n = n, mean = 0, sd = 1)
x2 <- rnorm(n = n, mean = 1, sd = 1)
x3 <- rnorm(n = n, mean = 2, sd = 1)
x4 <- rnorm(n = n, mean = 3, sd = 1)

# takes all samples except the first one
null_spec <- function(l, parameters) {
  purrr::map2(l, parameters, \(l, param) l - param)
}
stat_functions <- list(stat_anova_f)

# all means except first one
stat_assignments <- list(mu2 = 1, mu3 = 1, mu4 = 1)

pf <- PlausibilityFunction$new(
  null_spec = null_spec,
  stat_functions = stat_functions,
  stat_assignments = stat_assignments,
  x1, x2, x3, x4
)
pvalue2 <- pf$get_value(rep(0, 3))
```

```{r}
pvalue2
```

```{r, eval=FALSE}
means <- c(mean(x2) - mean(x1),
           mean(x3) - mean(x1),
           mean(x4) - mean(x1))
pf$set_point_estimate(means)
point_estimate2 <- pf$point_estimate
```

```{r}
point_estimate2
```

```{r, eval=FALSE}
pf$set_parameter_bounds(
  point_estimate = pf$point_estimate,
  conf_level = pf$max_conf_level
)

pf$set_grid(
  parameters = pf$parameters,
  npoints = 10
)
pf$evaluate_grid(grid = pf$grid)
grid2 <- pf$grid
```

```{r}
grid2 |> dplyr::filter(pvalue > 0.99)
```



